A tool set for random number generation on GPUs in R

### 2022-07-22 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3182 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

 <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE, tidy=FALSE>>=
# using GPU
 streams <- createStreamsGpu(n = 256*64)
 month_gpu <- vclMatrix(month, type = "integer")
 system.time(result_month <- 
   clrng::fisher.sim(month_gpu, 1e6, 
      streams=streams, type="double", 
      returnStatistics=TRUE,  Nglobal = c(256,64)))
 @
<<timecomparemonth1Gpuresult>>=
unlist(result_month[c('threshold','simNum','counts')])
result_month$p.value
@
We got 409429 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $\Sexpr{sn(result_month$p.value,3)}$, which is close enough to the p-value from \fct{stats::fisher.test}. \fct{clrng::fisher.sim} takes about 0.3 seconds, \pkg{clrng} accounts for 2.1\% of the elapsed time on CPU. 
 <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
 month_stats <- as.vector(result_month$sim)
 @

@@

Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
# dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv")
# week<-as.matrix(dataweek[,-1])
# rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
# colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
#week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
data("week")
as.matrix(week)
@
<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"),  #label = "tab:week",
             caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
kable_styling(full_width = F, position = "center")#, latex_options = "HOLD_position")
@
<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
# using GPU
week_GPU<-gpuR::vclMatrix(week,type="integer")
system.time(resultWeek<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
unlist(resultWeek[c('threshold','simNum','counts')])
resultWeek$p.value
@
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
resultWeek$cpu = as.vector(resultWeek$sim)
@
<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
# using CPU
system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
result_weekcpu$p.value
@
The ``week'' table has a much smaller p-value: around $\Sexpr{sn(result_weekcpu$p.value,3)}$, which should require a larger number of simulations to get a more accurate p-value. With more than ten million simulations, we get 1350 cases and a p-value around $\Sexpr{sn(resultWeek$p.value,3)}$ with \fct{fisher.sim}. \fct{stats::fisher.test} takes about 91.8 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.

>>>>

 <<timecomparemonth1Gpu, eval=TRUE,cache=FALSE, tidy=FALSE>>=
# using GPU
 streams <- createStreamsGpu(n = 256*64)
 month_gpu <- vclMatrix(month, type = "integer")
 system.time(result_month <- 
   clrng::fisher.sim(month_gpu, 1e6, 
      streams=streams, type="double", 
      returnStatistics=TRUE,  Nglobal = c(256,64)))
 @
<<timecomparemonth1Gpuresult>>=
unlist(result_month[c('p.value','threshold','simNum','counts')])
@
We got 409429 cases whose test statistics are below the observed threshold based on an actual number of 1015808 simulations, so the p-value from \fct{clrng::fisher.sim} for this table is about $\Sexpr{sn(result_month$p.value,3)}$, which is close enough to the p-value from \fct{stats::fisher.test}. \fct{clrng::fisher.sim} takes about 0.3 seconds, \pkg{clrng} accounts for 2.1\% of the elapsed time on CPU. 
 <<timecomparemonth3Gpu, eval=TRUE,echo=FALSE, cache=TRUE>>=
 month_stats <- as.vector(result_month$sim)
 @

@@

Table \ref{tab:week} is a $7 \times 12$ table shows frequencies for congenital anomalies of the newborn by birth day of week in 2018 within the United States. 
<<Timecompareweek, echo=FALSE, eval=TRUE, cache=TRUE,results="hide", fig.pos='h'>>=
# dataweek<-read.csv("/home/ruoyong/diseasemapping/pkg/gpuRandom/inst/documents/paper1_2021_5/weekday.csv")
# week<-as.matrix(dataweek[,-1])
# rownames(week) <- c("Mon", "Tue","Wed","Thu","Fri","Sat","Sun")
# colnames(week) <- c("Ane", "Men", "Cya", "Her", "Omp", "Gas", "Lim", "Cle", "Pal", "Dow", "Chr", "Hyp")
#week <- as.matrix(readRDS(system.file("extdata", "week.Rds", package = "clrng")))
data("week")
as.matrix(week)
@
<<weekdata, eval=TRUE, echo=FALSE, cache=TRUE>>=
knitr::kable(week,format="latex",align = c("rrrrrrrrrrrr"),  #label = "tab:week",
             caption = "Day-of-week birth anomaly data\\label{tab:week}", booktabs=TRUE,linesep = "") %>%
kable_styling(full_width = F, position = "center")#, latex_options = "HOLD_position")
@
<<TimecompareweekGpu2, eval=TRUE, cache=FALSE>>=
# using GPU
week_GPU<-gpuR::vclMatrix(week,type="integer")
system.time(resultWeek<-clrng::fisher.sim(week_GPU, 1e7, streams=streams,
                type="double",returnStatistics=TRUE,Nglobal = c(256,64)))
unlist(resultWeek[c('p.value','threshold','simNum','counts')])
@
<<TimecompareweekGpu3, echo=FALSE, eval=TRUE, cache=TRUE>>=
resultWeek$cpu = as.vector(resultWeek$sim)
@
<<TimecompareweekCpu, cache=TRUE, eval=TRUE>>=
# using CPU
system.time(result_weekcpu<-fisher.test(week,simulate.p.value = TRUE,B=10010624))
result_weekcpu$p.value
@
The ``week'' table has a much smaller p-value: around $\Sexpr{sn(result_weekcpu$p.value,3)}$, which should require a larger number of simulations to get a more accurate p-value. With more than ten million simulations, we get 1350 cases and a p-value around $\Sexpr{sn(resultWeek$p.value,3)}$ with \fct{fisher.sim}. \fct{stats::fisher.test} takes about 91.8 seconds, while \fct{fisher.sim} takes about 1.9 seconds, the GPU elapsed time is decreased to about 2.1\% of the time taken on CPU.
